\chapter{Background \& Objectives}

The aim of this project is to implement a UNIX command line shell that conforms to the POSIX shell standard\cite{POSIX-SHELL-STANDARD}.

A command line shell is the minimal interface between users and the operating system kernel and allows users to work with a relatively simple read-eval-print loop.

%This section should discuss your preparation for the project, including background reading, your analysis of the problem and the process or method you have followed to help structure your work.  It is likely that you will reuse part of your outline project specification, but at this point in the project you should have more to talk about. 

\textbf{Note}: 

\begin{itemize}
   \item All of the sections and text in this example are for illustration purposes. The main Chapters are a good starting point, but the content and actual sections that you include are likely to be different.
   \item Look at the document on the Structure of the Final Report for additional guidance. 
\end {itemize}

\section{Background}
I became interested in doing this project during my year in industry when the ShellShock\cite{SHELLSHOCK-CVE,SHELLSHOCK-LWN,SHELLSHOCK-SYMANTEC} vunerability was annouced.
Bash is the default shell on most linux distributions and is used by many applications including web servers such as apache to serve CGI scripts.
This means that there was a huge attack surface for this bug.

I started examining the source code of bash once the patches for shellshock and the subsequently discovered vunerabilities were released.
The first thing I noticed was how insanely confusing it all was.

\begin{table}[]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{@{}lllll@{}}
\toprule
Language           & files & blank & comment & code   \\ \midrule
C                  & 252   & 20090 & 19392   & 102608 \\
HTML               & 3     & 3762  & 39      & 25942  \\
Bourne Shell       & 36    & 3330  & 3459    & 19857  \\
Teamcenter def     & 44    & 2567  & 1491    & 13014  \\
C/C++ Header       & 111   & 2719  & 3438    & 7286   \\
yacc               & 2     & 800   & 885     & 5096   \\
m4                 & 4     & 465   & 438     & 4656   \\
Perl               & 2     & 535   & 834     & 4229   \\
Bourne Again Shell & 24    & 196   & 249     & 786    \\
make               & 3     & 48    & 36      & 110    \\
Assembly           & 2     & 11    & 20      & 48     \\
awk                & 1     & 8     & 15      & 24     \\
sed                & 2     & 0     & 0       & 16     \\ \midrule
SUM:               & 486   & 34531 & 30296   & 183672 \\ \bottomrule
\end{tabular}
\end{table}






I have used bash ever since I started programming and switched to using linux as my main operating system.
I often use it during code golf competitions due to its terse yet powerful syntax.
What was your background preparation for the project? What similar systems did you assess? What was your motivation and interest in this project? 

\section{Analysis}
Taking into account the problem and what you learned from the background work, what was your analysis of the problem? How did your analysis help to decompose the problem into the main tasks that you would undertake? Were there alternative approaches? Why did you choose one approach compared to the alternatives? 

There should be a clear statement of the objectives of the work, which you will evaluate at the end of the work. 

In most cases, the agreed objectives or requirements will be the result of a compromise between what would ideally have been produced and what was felt to be possible in the time available. A discussion of the process of arriving at the final list is usually appropriate.

\section{Process}
You need to describe briefly the life cycle model or research method that you used. You do not need to write about all of the different process models that you are aware of. Focus on the process model that you have used. It is possible that you needed to adapt an existing process model to suit your project; clearly identify what you used and how you adapted it for your needs.

